Computing astrophysical inferences with diffusive nested sampling

### 2014-03-06 ms.tex
Some words and refs

Bayesian inference is usually performed by
constructing a set of possible answers to the question of
interest, that is, the ``hypothesis space'', indexed by parameters $\pars$.
A state of prior knowledge about the parameters is described by assigning a
prior distribution $p(\pars)$. Some relevant data $\data$ is about to become
known, and prior knowledge about a connection between $\theta$ and $\data$ is
expressed by a choice of ``sampling distribution'' $p(\data | \theta)$, which
is really a family of probability distributions, one for each possible value of
$\theta$. By the product rule, the joint (prior) distribution is
\begin{eqnarray}
p(\pars, \data) = p(\pars) p(\data | \pars) = p(\data)p(\pars | \data)
\end{eqnarray}
When the
data $\data$ becomes known, our state of knowledge about $\pars$ changes
from the marginal distribution $p(\pars)$ to the conditional distribution
$p(\pars | \data_{\rm actual})$
that corresponds to the observed data. The distribution
$p(\pars | \data_{\rm actual})$ is
called the ``posterior''.
Incidentally, Bayesian inference can be understood as a special case of
MaxEnt updating, starting from the joint prior $p(\pars, \data)$ and applying
the constraint $P(\data = \data_{\rm actual}) = 1$ \citep{caticha}. Note that
this is not a ``derivation'' of Bayesian updating as the rules of probability
already imply a commitment to update from $p(\theta)$ to $p(\theta | \data)$
once $\data$ becomes known.

@@

The zeroth component of this mixture is just the prior $p(\pars)$. The first
mixture component is the prior, but restricted to the domain where the
likelihood is greater than a threshold $L^*_1$. The second mixture component
is the prior, but restricted to the domain where the likelihood is above
$L^*_2$. The likelihood thresholds $\{ L^*_i \}$ are chosen so each
successive constrained prior occupies approximately $\exp(-1)$ times as much
prior mass (``volume'') as the previous constrained prior. This is strongly
related to the algorithm known as ``simulated tempering''
\citep{simulated_tempering}, but uses NS's
sequence of intermediate distributions instead of an annealed sequence.
Retaining ``old'' distributions in the mixture allows

\section{Probabilistic Catalogs}
The hypothesis space is the set of possible answers to the question
``how many objects are there, and what are their properties (positions, fluxes,
etc)'' \citep{starfield}.

%\section{Reverberation Mapping}
%Many objects in the universe are too distant to resolve, so we cannot simply
%``see'' what is going on. Rather we must infer what is going on. Active Galactic
%Nuclei (AGN) are a good example.

@@

\begin{thebibliography}{9}

@@

>>>>

Bayesian inference is usually performed by
constructing a set of possible answers to the question of
interest, that is, the ``hypothesis space'', indexed by parameters $\pars$.
A state of prior knowledge about the parameters is described by assigning a
prior distribution $p(\pars)$. Some relevant data $\data$ is about to become
known, and prior knowledge about a connection between $\theta$ and $\data$ is
expressed by a choice of ``sampling distribution'' $p(\data | \theta)$, which
is really a family of probability distributions, one for each possible value of
$\theta$. By the product rule, the joint (prior) distribution is
\begin{eqnarray}
p(\pars, \data) = p(\pars) p(\data | \pars) = p(\data)p(\pars | \data)
\end{eqnarray}
When the
data $\data$ becomes known, our state of knowledge about $\pars$ changes
from the marginal distribution $p(\pars)$ to the conditional distribution
$p(\pars | \data_{\rm actual})$
that corresponds to the observed data. The distribution
$p(\pars | \data_{\rm actual})$ is
called the ``posterior''.
Incidentally, Bayesian inference can be understood as a special case of
MaxEnt updating, starting from the joint prior $p(\pars, \data)$ and applying
the constraint $P(\data = \data_{\rm actual}) = 1$ \citep{caticha}. Note that
this is not a ``derivation'' of Bayesian updating as the rules of probability
already imply a commitment to update from $p(\theta)$ to $p(\theta | \data)$
once $\data$ becomes known. For convenience, once the data set has been
established and attention turns to computation, the prior is denoted by
$\pi(\theta)$ and the likelihood function is $L(\theta)$.

@@

The zeroth component of this mixture is just the prior $p(\pars)$. The first
mixture component is the prior, but restricted to the domain where the
likelihood is greater than a threshold $L^*_1$. The second mixture component
is the prior, but restricted to the domain where the likelihood is above
$L^*_2$. The likelihood thresholds $\{ L^*_i \}$ are chosen so each
successive constrained prior occupies approximately $\exp(-1)$ times as much
prior mass (``volume'') as the previous constrained prior. This is strongly
related to the algorithm known as ``simulated tempering''
\citep{simulated_tempering}, but uses NS's
sequence of intermediate distributions instead of an annealed sequence
based on $\pi(\theta)L(\theta)^\beta$. Retaining ``old'' distributions in the
mixture gives the particles a chance to explore more freely, perhaps escaping
one mode in the distribution and moving to another.

A multi-threaded, free software C++ implementation of Diffusive Nested Sampling
is available online at {\tt http://github.com/eggplantbren/DNest3}.

\section{Probabilistic Catalogs}
Many questions in astronomy can be classified as ``source detection'' problems.
Telescopes (whether optical, radio, or whatever) obtain some data, and we would
like to use this data to infer the presence and properties of the objects in
that part of the sky. Several different techniques and software packages have
been developed for carrying this out \citep[e.g.][]{sextractor, dolphot}.
However, this process
can also be described as a Bayesian inference problem \citep{starfield}.
The hypothesis space is the set of possible answers to the question
``how many objects are there, and what are their properties (positions, fluxes,
etc)''. The prior

@@

\begin{thebibliography}{9}
\bibitem[Bertin and
Arnouts(1996)]{sextractor} Bertin, E., Arnouts, S.\ 1996.\ SExtractor: Software for source extraction.\ Astronomy and Astrophysics Supplement Series 117, 393-404.

@@

\bibitem[Dolphin(2000)]{dolphot} Dolphin, A.~E.\ 2000, Publications of the
Astronomical Society of the Pacific, 112, 1383
